---
title: "Block groups"
author: "Steve Simon and Satya Golla"
date: "2/7/2022"
output: html_document
---



```{r setup}
library(sf)
library(tidyverse)
```

### census block groups

A census block group is a statistical division of census tracts, and generally defined to contain between 600 and 3000 people. There are files for census block groups for [Kansas][c_bg_ks] and [Missouri][c_bg_mo]. 


```{r read-bg-ks}
path_name <- "./Data"
file_name <- "tl_2021_20_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> all_bg_ks
head(all_bg_ks)
all_bg_ks                              %>%
     ggplot()                           +
      geom_sf(aes())                    +
      ggtitle("ks-bg")
```


```{r read-bg-mo}
path_name <- "./Data"
file_name <- "tl_2021_29_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> all_bg_mo
head(all_bg_mo)
all_bg_mo                              %>%
     ggplot()                           +
      geom_sf(aes())                    +
      ggtitle("mo-bg")
```



```{r combine-bg} 
all_bg_ks                           %>%
  bind_rows(all_bg_mo)  -> mo_ks_bg 
mo_ks_bg                           %>%
 ggplot()                               +
      geom_sf(aes())                    +
      ggtitle("mo-ks-bg")  
```

### Community districts

This shapefile, Community_District, was provided by Neal Wilson. The various functions described in the US counties shapefiles section will be used here.


```{r read-Community-Districts}
file_name <- "Community_District.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> Community_Districts
head(Community_Districts)
Community_Districts                  %>%
  ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("Community_Districts")
Community_Districts %>%
  st_transform(crs=st_crs(mo_ks_bg)) -> cd
```

#### Intersection of block groups with community districts

Each community district will reside entirely inside a single county, and it is obvious to anyone familiar with the Kansas City metropolitan area which county is would be, so this is done mostly for the sake of continuity and completeness.

It is important that both the block groups and the community districts use a common coordinate reference system. It is unclear WHICH coordinate reference system is best, but this is controlled by the st_crs function of the sf package.

The st_intersects function will evaluate to TRUE if two shapefiles intersect. The st_area function will give the area of a shapefile and the st_intersection function will produce a new shapefile from the intersection of two shapefiles.


```{r intersect-mo-ks-bg-Community-Districts}
for (i in 1:2) { 
mo_ks_bg                      %>%
  st_intersects( cd[i,], sparse=FALSE) -> intersect_matrix
  total_area  <- st_area(mo_ks_bg[intersect_matrix,]) / 1000000
  intersect_area <- st_area(st_intersection(mo_ks_bg[intersect_matrix,],cd[i,])) / 1000000 
  print(intersect_area/total_area)
 mo_ks_bg[intersect_matrix,]               %>%
 ggplot(aes())                                  +
        geom_sf(fill=NA, col="darkgreen") +
        geom_sf(data=cd[i, ], fill=NA, col="darkred") +
        ggtitle("mo-ks-tracts-Community-Districts") -> map
  print(map)}
```


