---
title: "Counties"
author: "Steve Simon and Satya Golla"
date: "12/29/2021"
output: html_document
---

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Miami County, KS (20121)
Wyandotte County, KS (20209)

```{r setup}
library(sf)
library(tidyverse)
```

```{r read-county}
path_name <- "./data"
file_name <- "tl_2021_us_county.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> all_counties
head(all_counties)
all_counties                           %>%
   filter(STATEFP==20)                    %>% 
   ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("ks-counties")
all_counties                           %>%
   filter(STATEFP==29)                    %>% 
   ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("mo-counties")
all_counties                           %>%
   filter(STATEFP==20|STATEFP==29)     %>% 
   ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("mo-ks-counties")
```



variable in the file 

+ Variables:
  +  "STATEFP"
  +   "COUNTYFP"
  +   "TRACTCE"
  +   "BLKGRPCE" 
  +   "AFFGEOID" 
  +   "GEOID"
  +   "NAME"
  +   "NAMELSAD"
  +   "LSAD"
  +   "ALAND"
  +   "AWATER"
  +   "geometry"
 

### Community districts

This shapefile, Community_District, was provided by Neal Wilson. The various functions described in the US counties shapefiles section will be used here.

```{r read-Community-Districts}
file_name <- "Community_District.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> Community_Districts
head(Community_Districts)
Community_Districts                  %>%
  ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("Community_Districts")
Community_Districts %>%
  st_transform(crs=st_crs(all_counties)) -> cd
```


### Neighborhoods

This shapefile, Neighborhoods, was provided by Neal Wilson. The various functions described in the US counties shapefiles section will be used here.


```{r read-Neighborhoods}
file_name <- "Neighborhoods.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> Neighborhoods
head(Neighborhoods)
Neighborhoods                          %>%
  ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("Neighborhoods")
Neighborhoods                          %>%
  st_transform(crs=st_crs(all_counties)) -> nbd
```

### Calculate intersections

#### Intersection of counties with community districts

Each community district will reside entirely inside a single county, and it is obvious to anyone familiar with the Kansas City metropolitan area which county is would be, so this is done mostly for the sake of continuity and completeness.

It is important that both the counties and the community districts use a common coordinate reference system. It is uncelar WHICH coordinate reference system is best, but this is controlled by the st_crs function of the sf package.

The st_intersects function will evaluate to TRUE if two shapefiles intersect. The st_area function will give the area of a shapefile and the st_intersection function will produce a new shapefile from the intersection of two shapefiles.

```{r intersect-jackson-county-Community-Districts}
all_counties                         %>%
  filter(STATEFP==20 | STATEFP==29)  ->       ks_mo_counties
ks_mo_counties                       %>%
  filter(STATEFP==29 & COUNTYFP=="095")               %>%
  st_intersects( cd, sparse=FALSE) -> intersect_matrix
 print(intersect_matrix)
 ggplot(data=cd, aes()) +
        geom_sf(fill="lightgreen", col="darkgreen") +
        geom_sf(data=cd[intersect_matrix, ], fill="pink", col="darkred") +
        ggtitle("jackson-county-Community-Districts")
```


```{r intersection-area-jackson-county}
total_area  <- st_area(cd[intersect_matrix,]) / 1000000
total_area
ks_mo_counties  %>%
  filter(STATEFP==29 & COUNTYFP=="095")  -> jackson_county
intersect_area <- st_area(st_intersection(cd[intersect_matrix,],jackson_county)) / 1000000
intersect_area
pct_area  <- round(intersect_area/total_area * 100, 1)
pct_area
st_centroid(cd[intersect_matrix,])
cd[1,]
```


```{r map-percentages}
 ggplot(data=cd, aes()) +
        geom_sf(fill="lightgreen", col="darkgreen") +
        geom_sf(data=cd[intersect_matrix,][as.numeric(pct_area)>10,], fill="pink", col="darkred") +
        ggtitle("jackson-county-Community-Districts") 
```


#### Intersection of counties with neighborhoods

Again, each neighborhood will reside in a single county. The various functions described in the intersection of counties with community districts section will be used here. 


```{r intersect-counties-Neighborhoods}
all_counties                         %>%
  filter(STATEFP==20 | STATEFP==29)  ->       ks_mo_counties
ks_mo_counties                       %>%
  filter(STATEFP==29 & COUNTYFP=="095")               %>%
  st_intersects( nbd, sparse=FALSE) -> intersect_matrix
 print(intersect_matrix)
 ggplot(data=nbd, aes()) +
        geom_sf(fill="lightgreen", col="darkgreen") +
        geom_sf(data=nbd[intersect_matrix, ], fill="pink", col="darkred") 
```
