---
title: "PUMAS"
author: "Steve Simon and Satya Golla"
date: "12/30/2021"
output: html_document
---

```{r setup}
library(sf)
library(tidyverse)
```

### Public Use Microdata Areas (PUMA)

A PUMA is a region within a state that contains no less than 100,000 people. There are files for PUMAs for [Kansas][c_puma_ks] and [Missouri][c_puma_mo]. The various functions described in the US counties shapefiles section will be used here.



```{r read-puma-ks}
path_name <- "../Data"
file_name <- "tl_2021_20_puma10.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> all_puma_ks
head(all_puma_ks)
all_puma_ks                           %>%
     ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("ks-puma")
```

variable in the file 

+ Variables:
  +   "STATEFP10"
  +   "PUMACE10"
  +   "GEOID10"
  +   "NAMELSAD10"
  +   "MTFCC10" 
  +   "FUNCSTAT10" 
  +   "ALAND10"
  +   "AWATER10"
  +   "INTPTLAT10"
  +   "INTPTLON10"
  +   "geometry"


```{r read-puma-mo}
path_name <- "../data"
file_name <- "tl_2021_29_puma10.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> all_puma_mo
head(all_puma_mo)
all_puma_mo                           %>%
     ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("mo-puma")
```


variable in the file 

+ Variables:
  +   "STATEFP10"
  +   "PUMACE10"
  +   "GEOID10"
  +   "NAMELSAD10"
  +   "MTFCC10" 
  +   "FUNCSTAT10" 
  +   "ALAND10"
  +   "AWATER10"
  +   "INTPTLAT10"
  +   "INTPTLON10"
  +   "geometry"

```{r combine-pumas}
all_puma_ks                %>%
  bind_rows(all_puma_mo)  -> mo_ks_puma 
mo_ks_puma %>%
 ggplot()                               +
      geom_sf(aes())                    +
      ggtitle("mo-ks-puma")  
```


### Community districts

This shapefile, Community_District, was provided by Neal Wilson. The various functions described in the US counties shapefiles section will be used here.

```{r read-Community-Districts}
file_name <- "Community_District.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> Community_Districts
head(Community_Districts)
Community_Districts                  %>%
  ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("Community_Districts")
Community_Districts %>%
  st_transform(crs=st_crs(mo_ks_puma)) -> cd
```

#### Intersection of pumas with community districts

Each community district will reside entirely inside a single county, and it is obvious to anyone familiar with the Kansas City metropolitan area which county is would be, so this is done mostly for the sake of continuity and completeness.

It is important that both the pumas and the community districts use a common coordinate reference system. It is unclear WHICH coordinate reference system is best, but this is controlled by the st_crs function of the sf package.

The st_intersects function will evaluate to TRUE if two shapefiles intersect. The st_area function will give the area of a shapefile and the st_intersection function will produce a new shapefile from the intersection of two shapefiles.


```{r intersect-puma-Community-Districts}
for(i in 1:2) {
mo_ks_puma                             %>%
st_intersects( cd[i,], sparse=FALSE) -> intersect_matrix
  total_area  <- st_area(mo_ks_puma[intersect_matrix,]) / 1000000
  intersect_area <- st_area(st_intersection(mo_ks_puma[intersect_matrix,],cd[i,])) / 1000000 
  print(intersect_area/total_area)
 map_position <- st_centroid(mo_ks_puma[intersect_matrix, ])
 ggplot(data=mo_ks_puma[intersect_matrix, ], aes()) +
        geom_sf(fill=NA, col="darkgreen") +
        geom_sf(data=cd[i,], fill=NA, col="darkred")+
   ggtitle(paste(c(cd$NAME[i],mo_ks_puma$GEOID10[intersect_matrix]),collapse = ", ")) -> map
   print(map)
}
```


### Neighborhoods

This shapefile, Neighborhoods, was provided by Neal Wilson. The various functions described in the US counties shapefiles section will be used here.

```{r read-Neighborhoods, eval = FALSE}
file_name <- "Neighborhoods.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      -> Neighborhoods
head(Neighborhoods)
Neighborhoods                          %>%
  ggplot()                              +
      geom_sf(aes())                    +
      ggtitle("Neighborhoods")
Neighborhoods                          %>%
  st_transform(crs=st_crs(mo_ks_puma)) -> nbd
```


```{r intersect-puma-Neighborhoods}
for(i in 1:2) {
mo_ks_puma               %>%
 st_transform(crs=st_crs(mo_ks_puma)) -> nbd
 st_intersects( nbd, sparse=FALSE) -> intersect_matrix
 print(intersect_matrix)
 ggplot(data=nbd, aes()) +
        geom_sf(fill="lightgreen", col="darkgreen") +
        geom_sf(data=nbd[intersect_matrix, ], fill="pink", col="darkred") 
} 
```
