
---
title: "Calculate intersection of blocks and community districts"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-11"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---

This program will calculate which census blocks intersect with which community districts. It relies on information stored in

+ bg.RData
+ bl.RData
+ cd.RData
+ cd-intersections.RData
+ red.RData

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Counts

There are 7 counties in the Kansas City metropolitan area.

There are 59 community districts in the 7 counties.

There are 1,684 block groups in the 7 counties. 776 of these intersect partially or completely with one or more community districts. 465 intersect with only one community district, 213 with two, 65 with 3, 21 with 4, 7 with 5, 4 with 6, and 1 with 7. Note that `465*1 + 213*2 +65*3 + 21*4 + 7*5 + 4*6 + 1*7 = 1,236`.

There are 39,903 blocks in the 7 counties. 17,547 of these intersect partially or completely with one or more community districts. 16339 intersect with only one community district, 1,108 intersect with two, 91 intersect with three, and 9 intersect with four. Note that `16339*1 + 1108*2 + 91*3 + 9*4 = 18864`.

### Setup

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
```

### Load key files

Block groups

```{r load-bg}
load(paste0(path_name, "bg", ".RData"))
glimpse(bg)
```

Blocks

```{r load-bl}
load(paste0(path_name, "bl", ".RData"))
glimpse(bl)
```

Community districts

```{r load-cd}
load(paste0(path_name, "cd", ".RData"))
glimpse(cd)
```

Community district intersections

```{r load-cd-intersections}
load(paste0(path_name, "cd-intersections", ".RData"))
glimpse(bg_cd_intersection)
glimpse(bl_cd_intersection)
```

```{r load-redistricting}
load(paste0(path_name, "red", ".RData"))
glimpse(red)
```

### Get block group list

Identify the list of all block groups that partially or completely intersect with any community district.

```{r intersect-bg}
bg_cd_intersection                     %>%
  pull(bg_id)                          %>%
  unique                                -> bg_list
glimpse(bg_list)
```

### Get block list

Now get a list of all census blocks that lie inside any of these block groups. Merge with red to get population and housing unit counts.

```{r pull-blocks}
bl                                     %>%
  pull(bl_id)                          %>%
  str_sub(1, 12)                       %>%
  bind_cols(bl$bl_id)                  %>%
  set_names(c("bg_id", "bl_id"))       %>%
  filter(bg_id %in% bg_list)           %>%
  inner_join(red, 
    by=c("bg_id", "bl_id"))               -> bl_list
glimpse(bl_list)
```


### Calculate counts and areas of intersections

```{r calculate-proportions}
bl_cd_intersection                     %>%
  full_join(bl_list, 
    by=c("bg_id", "bl_id"))            %>%
  replace_na(list(bl_prop_in=0))       %>%
  replace_na(list(cd_id=0))            %>%
  mutate(
    people_in=
      round(people*bl_prop_in))        %>%
  mutate(
    units_in=
      round(units*bl_prop_in))         %>%
  select(
    bg_id,
    bl_id,
    cd_id,
    bl_prop_in,
    people,
    people_in,
    units,
    units_in)                           -> bl_counts

glimpse(bl_counts)
```

### Calculate counts for block groups

```{r count-bl}
red                                    %>%
  group_by(bg_id)                      %>%
  summarize(
    people=sum(people),
    units=sum(units))                   -> count_total
glimpse(count_total)

bl_counts                              %>%
  group_by(bg_id, cd_id)               %>%
  summarize(
    people_in=sum(people_in),
    units_in=sum(units_in))            %>%
  filter(cd_id > 0)                     -> count_in
glimpse(count_in)

count_in                               %>%
  inner_join(count_total, by="bg_id")  %>%
  filter(cd_id > 0)                     -> bg_counts
glimpse(bg_counts)
```

### Save the information in an RData file.

```{r save}
save(
  bg_counts,
  bg_list,
  bl_counts,
  bl_list,
  file="../Data/cd-weights.RData")
```