
---
title: "Calculate intersection of blocks and community districts"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-11"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---

This program will calculate which census blocks intersect with which community districts. It relies on information stored in

+ bg.RData
+ bl.RData
+ cd.RData
+ cd-intersections.RData
+ red.RData

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Setup

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
```

### Load key files

Block groups

```{r load-bg}
load(paste0(path_name, "bg", ".RData"))
glimpse(bg)
```

Blocks

```{r load-bl}
load(paste0(path_name, "bl", ".RData"))
glimpse(bl)
```

Community districts

```{r load-cd}
load(paste0(path_name, "cd", ".RData"))
glimpse(cd)
```

Community district intersections

```{r load-cd-intersections}
load(paste0(path_name, "cd-intersections", ".RData"))
glimpse(bg_cd_intersection)
glimpse(bl_cd_intersection)
```

```{r load-redistricting}
load(paste0(path_name, "red", ".RData"))
glimpse(red)
```

### Get block group list

Identify the list of all block groups that partially or completely intersect with any community district.

```{r intersect-bg}
bg_cd_intersection                     %>%
  pull(bg_id)                          %>%
  unique                                -> bg_list
glimpse(bg_list)
```

### Get block list

Now get a list of all census blocks that lie inside any of these block groups.

```{r merge-blocks}
bl                                     %>%
  pull(bl_id)                          %>%
  str_sub(1, 12)                       %>%
  bind_cols(bl$bl_id)                  %>%
  set_names(c("bg_id", "bl_id"))       %>%
  filter(bg_id %in% bg_list)            -> bl_list
glimpse(bl_list)
```

### Calculate counts and areas of intersections

```{r calculate-proportions}
bl_cd_intersection                     %>%
  st_area                              %>%
  as.numeric                           %>%
  tibble                               %>%
  set_names("in_area")                 %>%
  bind_cols(bl_cd_intersection)        %>%
  mutate(bg_id=str_sub(bl_id, 1, 12))  %>%
  mutate(prop_in=in_area/bl_area)      %>%
  select(
    bg_id,
    bl_id, 
    cd_id, 
    prop_in)                           %>%
  full_join(bl_list, 
    by=c("bg_id", "bl_id"))            %>%
  replace_na(list(prop_in=0))          %>%
  full_join(red, 
    by=c("bg_id", "bl_id"))            %>%
  mutate(
    people_in=round(people*prop_in))   %>%
  mutate(
    units_in=round(units*prop_in))     %>%
  select(
    bg_id,
    bl_id,
    cd_id,
    prop_in,
    people_in,
    units_in)                           -> bl_counts

glimpse(bl_counts)
```

### Calculate counts for block groups

```{r count-bl}
red                                    %>%
  filter(bg_id %in% bg_list)           %>%
  group_by(bg_id)                      %>%
  summarize(
    people=sum(people),
    units=sum(units))                   -> count_total

bl_counts                              %>%
  group_by(bg_id, cd_id)               %>%
  summarize(
    people_in=sum(people_in),
    units_in=sum(units_in))             -> count_in

count_in                               %>%
  inner_join(count_total, by="bg_id")  %>%
  filter(!is.na(cd_id))                 -> bg_counts
glimpse(bg_counts)
```

### Save the information in an RData file.

```{r save}
save(
  bg_counts,
  bg_list,
  bl_counts,
  bl_list,
  file="../Data/cd-weights.RData")
```