---
title: "Cumulate people and houses across tracts"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-14"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---

This program will calculate the total number of people and houses within a census tract and the intersection with community districts.

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
load(paste0(path_name, "red", ".RData"))
load(paste0(path_name, "bl-cd", ".RData"))
```


### Summarize first by tracts

```{r summarize1}
red                                    %>%
  mutate(POP100=as.numeric(POP100))    %>%
  mutate(HU100=as.numeric(HU100))       -> red

red                                    %>%
  group_by(
    statefp, 
    countyfp, 
    tract)                             %>%
  summarize(
    tract_people=sum(POP100),
    tract_houses=sum(HU100),
    tract_blocks=n())                   -> tract_summary
glimpse(tract_summary)
```

### Then summarize by tracts and communities

```{r summarize2}
bl_cd                                  %>%
  mutate(POP100=as.numeric(POP100))    %>%
  mutate(HU100=as.numeric(HU100))      %>%
  group_by(
    statefp, 
    countyfp, 
    tract,
    cd_id)                             %>%
  summarize(
    cd_people=sum(POP100*pct_in/100),
    cd_houses=sum(HU100*pct_in/100),
    cd_blocks=n())                     %>%
    arrange(cd_id, tract)               -> cd_summary
glimpse(cd_summary)
```

### Combine the two summaries

```{r combine}
cd_summary                              %>%
  full_join(
    tract_summary, 
    by=c(
      "statefp",
      "countyfp",
      "tract"))                         %>%
  mutate(w1=cd_people/tract_people)     %>%
  mutate(w2=cd_houses/tract_houses)      -> cd_w
glimpse(cd_w)
```

### Store the result

```{r store}
save(cd_w, file="../Data/cd-w.RData")