---
title: "Calculate intersection of blocks and community districts"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-11"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })
---

This program will calculate which census blocks intersect with which community districts. It relies on information stored in

+ bl.RData
+ cd.RData
+ red.RData

### General information

For information about various files, refer to the files that stored the relevant data (e.g., store_block_information.Rmd).

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Wyandotte County, KS (20209)

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
```

### Revise and simplify redistricting information

```{r revise-red}
load(paste0(path_name, "red", ".RData"))
red                                      %>%
  select(-GEOID)                          -> rev_red
```

### Revise and simplify block information

```{r revise-block}
load(paste0(path_name, "bl", ".RData"))
bl                                       %>%
  select(
    statefp,
    countyfp,
    tract,
    block,
    block_area,
    geometry)                            %>%
  full_join(rev_red,
    by=c(
      "statefp",
      "countyfp",
      "tract",
      "block"))                           -> rev_bl
glimpse(rev_bl)
```

### Revise and simplify community information

```{r revise-cd}
load(paste0(path_name, "cd", ".RData"))
cd                                       %>%
  rename(cd_id=CD_ID)                    %>%
  select(
    cd_id,
    geometry)                            %>%
  mutate(
    cd_area=
      as.numeric(st_area(cd)))            -> rev_cd
glimpse(rev_cd)
```

### Intersect blocks and community districts

```{r intersect}
rev_bl                                   %>%
  st_intersection(rev_cd)                 -> bl_cd
bl_cd                                    %>%
  mutate(
    in_area=
      as.numeric(st_area(bl_cd)))        %>%
  mutate(
    pct_in=
      100*in_area/block_area)             -> bl_cd
glimpse(bl_cd)
```

### Save the information in an RData file.

```{r save}
save(bl_cd, file="../Data/bl-cd.RData")
```