---
title: "Map overlapping shapefiles"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-11"
output: html_document
---

This program will compare various shapefiles for counties, public use microdata areas, census tracts, census block groups, census blocks, zip code tabulation areas, community districts, and neighborhoods. It will produce a map showing overlaps of different shapefiles. In addition, the program will compute the area of each shapefile and what percentage of that area lies inside or outside another shapefile.

### General information

The starting point for downloading most Census shapefiles is [here][c_general]. The ftp site is a bit more difficult to navigate, but leads you directly to the files that you want. 

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Platte County, MO (29165)
Wyandotte County, KS (20209)

The community district and neighborhood shapefiles are documented in an email from Neal Wilson dated 2022-03-03. You can find a pdf file of that email in the doc folder.

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
```

```{r functions}
map_overlap <- function(shape1, shape2) {
  sb <- st_intersects(shape1, shape2, sparse=FALSE)

  shape2                              %>%
    filter(as.vector(sb))              -> shape2_sb

  names <- paste0(c(shape1$label, shape2_sb$label), collapse=", ")
  shape2_in <- st_intersection(shape1, shape2_sb)

  shape1                                %>%
    st_area                             %>%
    divide_by(1e6)                      %>%
    as.numeric                           -> a1

  shape2_sb                             %>%
    st_area                             %>%
    divide_by(1e6)                      %>%
    as.numeric                           -> a2
  
  shape2_in                             %>%
    st_area                             %>%
    divide_by(1e6)                      %>%
    as.numeric                           -> a2_in
  
  shape2_sb$label                       %>%
    data.frame                          %>%
    mutate(area1=round(a1, 2))          %>%
    mutate(area2=round(a2, 2))          %>%
    mutate(area2_in=round(a2_in, 2))    %>%
    mutate(pct1=round(100*a2_in/a1, 2)) %>%
    mutate(pct2=round(100*a2_in/a2, 2)) %>%
    print

  fill_color <- case_when(
    a2_in/a2 > 0.99~"pink",
    a2_in/a2 < 0.01~"lightgreen",
    TRUE~"lightblue")
  shape2_sb %>%
    ggplot(aes())                        +
    geom_sf(fill=fill_color, col="darkgreen")    +
    ggtitle(names)                       +
    geom_sf(
      data=shape1,
      fill=NA,
      size=1.2,
      col="darkred")                  
}
```

### Intersect counties with community districts

```{r county-cd}
load(paste0(path_name, "co", ".RData"))
load(paste0(path_name, "cd", ".RData"))
n <- 1:dim(cd)[1]
n <- 3
for (i in n) {
  map_overlap(cd[i, ], county) %>% print
}
```

### Intersect neighborhoods with community districts

```{r nbd-cd}
load(paste0(path_name, "nbd", ".RData"))
n <- 1:dim(cd)[1]
n <- 3
for (i in n) {
  map_overlap(cd[i, ], nbd) %>% print
}
```

### Intersect community districts with census blocks

```{r cd-block}
load(paste0(path_name, "bl", ".RData"))
n <- 1:dim(cd)[1]
n <- 1:3
for (i in n) {
  map_overlap(cd[i, ], bl) %>% print
}
```

<!---Key census links go here--->

[r_sf]: https://r-spatial.github.io/sf/

[r read_acs]: https://cran.r-project.org/package=acs

[c_general]: https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html

[c_counties]: https://www2.census.gov/geo/tiger/TIGER2021/COUNTY/tl_2021_us_county.zip
[c_puma_ks]: https://www2.census.gov/geo/tiger/TIGER2021/PUMA/tl_2021_20_puma10.zip
[c_puma_mo]:https://www2.census.gov/geo/tiger/TIGER2021/PUMA/tl_2021_29_puma10.zip
[c_tract_ks]: https://www2.census.gov/geo/tiger/TIGER2021/TRACT/tl_2021_20_tract.zip
[c_tract_mo]: https://www2.census.gov/geo/tiger/TIGER2021/TRACT/tl_2021_29_tract.zip
[c_bg]: https://www2.census.gov/geo/tiger/TIGER2021/BG/
[c_bg_ks]: https://www2.census.gov/geo/tiger/TIGER2021/BG/tl_2021_20_bg.zip
[c_bg_mo]: https://www2.census.gov/geo/tiger/TIGER2021/BG/tl_2021_29_bg.zip

Insert links for census block groups and census blocks.

[c_redistricting]: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html

[c_help]: https://www.census.gov/data/academy/courses/2020-census-redistricting-data.html

[c_redistricting_ks]: https://www2.census.gov/programs-surveys/decennial/2020/data/01-Redistricting_File--PL_94-171/Kansas/ks2020.pl.zip
[c_redistricting_mo]: https://www2.census.gov/programs-surveys/decennial/2020/data/01-Redistricting_File--PL_94-171/Missouri/mo2020.pl.zip

[c_r_code]: https://www2.census.gov/programs-surveys/decennial/rdo/about/2020-census-program/Phase3/SupportMaterials/2020PL_R_import_scripts.zip