---
title: "Quality check 01, Mismatches"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-14"
output: html_document
---

This program examines blocks that only partially intersect with community districts.

### General information

The starting point for downloading most Census shapefiles is [here][c_general]. You can find more about the census redistricting files [here][c_redistricting]. 

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Platte County, MO (29165)
Wyandotte County, KS (20209)

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
load(paste0(path_name, "bl", ".RData"))
glimpse(bl)
load(paste0(path_name, "cd", ".RData"))
glimpse(cd)
load(paste0(path_name, "icd", ".RData"))
glimpse(icd)
```

### List blocks that are split between communities

```{r list}
icd                                    %>%
  mutate(dev=abs(pct_in-50))           %>%
  arrange(dev)                          -> dev
dev                                    %>%
  print(n=20)
```

### Map split blocks

```{r map}
for (i in 1:500) {
  cd                                   %>%
    filter(cd_id==dev$cd_id[i])         -> cd_i
  bl                                   %>%
    filter(bl_id==dev$bl_id[i])         -> bl_i
  ti <- paste0(
    cd_i$CD_NAME, ", ",
    bl_i$NAME20, ", ",
    round(dev$pct_in[i]), "%")
  cd_i                                 %>%
    ggplot(aes())                       +
      geom_sf(
        fill="white", 
        color="darkgreen")              +
      ggtitle(ti)                       +
      geom_sf(
        data=bl_i,
        aes(), 
        fill=NA,
        color="darkred")                -> map_i
  plot(map_i)
}
```

<!---Key census links go here--->

[c_general]: https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html

[c_redistricting]: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html
