---
title: "Quality check 01, Mismatches"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-14"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program examines mismatches between census blocks shape files and redistricting files.

### General information

The starting point for downloading most Census shapefiles is [here][c_general]. You can find more about the census redistricting files [here][c_redistricting]. 

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Platte County, MO (29165)
Wyandotte County, KS (20209)

### Load relevant files

```{r setup}
library(sf)
library(tidyverse)
library(magrittr)
path_name <- "../data/"
load(paste0(path_name, "bl", ".RData"))
glimpse(bl)
load(paste0(path_name, "red", ".RData"))
glimpse(red)
```

### Calculate mismatches

```{r mismatch}
bl                                    %>%
  full_join(red, by=c("bl_id"))       %>%
  mutate(
    m_red=
      as.numeric(is.na(people)))      %>%
  mutate(
    m_bl=as.numeric(is.na(bl_name)))  %>%
  mutate(m_code=m_red+2*m_bl)          -> m
table(m$m_code, m$st_id)
table(m$m_code, m$co_id)
m                                     %>%
  ggplot(aes()) +
    geom_sf()
m                                     %>%
  filter(m_red==1)                    %>%
  ggplot(aes()) +
    geom_sf()
m                                     %>%
  filter(m_red==0)                    %>%
  ggplot(aes()) +
    geom_sf()
```

<!---Key census links go here--->

[c_general]: https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html

[c_redistricting]: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html

