---
title: "Quality check 04, Summing block data to block group level"
author: "Satya Golla and Steve Simon"
date: "Created 2022-04-02"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

Redistricing data has information about Census blocks (SUMLEV=750), block groups (SUMLEV=150), Census tracts (SUMLEV=140) and counties (SUMLEV=050).

Part 1: Do the POP100 and HU100 values sum up properly from the lowest level of the Census hierarchy to higher levels?

Part 2: Do the totals computed in calculate-cd-weights.Rmd match the numbers from the Census?

Part 3: Do the totals computed in calculate-cd-weights approximately match the numbers from the ACS?

For more information, see [this pdf file][sumlev].

## Preliminary code

### Setup

```{r setup}
library(sf)
library(tidycensus)
library(tidyverse)
path_name <- "../data"
vnames <- c(
  "FILEID", 
  "STUSAB", 
  "SUMLEV", 
  "GEOVAR", 
  "GEOCOMP", 
  "CHARITER", 
  "CIFSN", 
  "LOGRECNO", 
  "GEOID", 
  "GEOCODE",
  "REGION",
  "DIVISION",
  "STATE",
  "STATENS",
  "COUNTY",
  "COUNTYCC",
  "COUNTYNS",
  "COUSUB",
  "COUSUBCC",
  "COUSUBNS",
  "SUBMCD",
  "SUBMCDCC",
  "SUBMCDNS",
  "ESTATE",
  "ESTATECC",
  "ESTATENS", 
  "CONCIT",
  "CONCITCC",
  "CONCITNS",
  "PLACE",
  "PLACECC",
  "PLACENS",
  "TRACT",
  "BLKGRP",
  "BLOCK", 
  "AIANHH",
  "AIHHTLI",
  "AIANHHFP",
  "AIANHHCC",
  "AIANHHNS",
  "AITS",
  "AITSFP",
  "AITSCC",
  "AITSNS",
  "TTRACT",
  "TBLKGRP",
  "ANRC",
  "ANRCCC",
  "ANRCNS",
  "CBSA",
  "MEMI",
  "CSA",
  "METDIV",
  "NECTA",
  "NMEMI",
  "CNECTA",
  "NECTADIV",
  "CBSAPCI",
  "NECTAPCI",
  "UA",
  "UATYPE",
  "UR",
  "CD116",
  "CD118",
  "CD119", 
  "CD120", 
  "CD121", 
  "SLDU18", 
  "SLDU22", 
  "SLDU24", 
  "SLDU26", 
  "SLDU28", 
  "SLDL18", 
  "SLDL22",
  "SLDL24",
  "SLDL26",
  "SLDL28",
  "VTD",
  "VTDI",
  "ZCTA",
  "SDELM",
  "SDSEC",
  "SDUNI",
  "PUMA",
  "AREALAND",
  "AREAWATR",
  "BASENAME",
  "NAME",
  "FUNCSTAT",
  "GCUNI",
  "POP100",
  "HU100",
  "INTPTLAT",
  "INTPTLON",
  "LSADC",
  "PARTFLAG",
  "UGA")
```

### Read Kansas redistricting files

```{r read-ks}
file_name <- "red/ksgeo2020.pl"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  read.delim(
    header=FALSE, 
    colClasses="character", 
    sep="|")                           -> red1
names(red1) <- vnames
```

### Read Missouri redistricting files

```{r read-mo}
file_name <- "red/mogeo2020.pl"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  read.delim(
    header=FALSE, 
    colClasses="character", 
    sep="|")                           -> red2
names(red2) <- vnames
```

## Part 1.

### Combine and extract block information

```{r combine}
red1                                   %>%
  bind_rows(red2)                      %>%
  mutate(POP100=as.numeric(POP100))    %>%
  mutate(HU100=as.numeric(HU100))       -> red

red                                    %>%
  filter(SUMLEV=="750")                %>%
  rename(bl_id=GEOCODE)                %>%
  rename(bl_people=POP100)             %>%
  rename(bl_units=HU100)       %>%
  mutate(bg_id=str_sub(bl_id, 1, 12))  %>%
  select(
    bl_id, 
    bg_id, 
    bl_people, 
    bl_units)                           -> red_bl
```

### Extract block group information

```{r extract-bg}
red                                    %>%
  filter(SUMLEV=="150")                %>%
  rename(bg_id=GEOCODE)                %>%
  rename(bg_people=POP100)             %>%
  rename(bg_units=HU100)               %>%
  select(
    bg_id, 
    bg_people, 
    bg_units)                           -> red_bg
```

### Compare block totals to block groups

```{r compare}
red_bl                                 %>%
  group_by(bg_id)                      %>%
  summarize(
    bl_people=sum(bl_people),
    bl_units=
      sum(bl_units))                   %>%
  full_join(red_bg, by="bg_id")        %>%
  mutate(
    diff_people=bl_people-bg_people)   %>%
  mutate(
    diff_units=
      bl_units - 
      bg_units)                         -> red_diff
head(red_diff[ , -1])
table(red_diff$diff_people)
table(red_diff$diff_units)
```

## Part 2

### Read cd-weights.RData

```{r cd-weights}
load("../Data/cd-weights.RData")
glimpse(bg_counts)
glimpse(bg_list)
glimpse(bl_counts)
glimpse(bl_list)
```

### Compare to redistricting files

```{r compare-bg-counts}
head(red_bg)
head(bg_counts)
red_bg                                 %>%
  inner_join(bg_counts, by="bg_id")    %>%
  mutate(diff_people=bg_people-people) %>%
  mutate(diff_units=bg_units-units)     -> diff_bg
head(diff_bg[ , -1])
table(diff_bg$diff_people)
table(diff_bg$diff_units)
```

## Part 3.

### Get ACS data

+ Key counties
  + Johnson County, KS (20091)
  + Leavenworth County, KS (20103)
  + Wyandotte County, KS (20209)
  + Cass County, MO (29037)
  + Clay County, MO (29047)
  + Jackson County, MO (29095)
  + Platte County, MO (29165)

```{r get-acs-units}
units_list <- c(
  "B25001_001",
  "B25002_001",
  "B25002_002",
  "B25002_003",
  "B25003_001",
  "B25004_001")
county_list <- c(
  "20091",
  "20103",
  "20209",
  "29037",
  "29047",
  "29095",
  "29165")
acs1 <- get_acs(
  geography="block group", 
  variables="B25002_001",
  state="KS",
  county=c("91","103", "209"),
  geometry=FALSE, 
  cache_table=TRUE,
  year=2020)
glimpse(acs1)

acs2 <- get_acs(
  geography="block group", 
  variables="B25002_001",
  state="MO",
  county=c("37","47", "95", "165"),
  geometry=FALSE, 
  cache_table=TRUE,
  year=2020)
glimpse(acs2)

acs1                                   %>%
  bind_rows(acs2)                      %>%
  rename(bg_id=GEOID)                  %>%
  select(-NAME)                        %>%
  inner_join(bg_counts, by="bg_id")    %>%
  select(
    bg_id,
    variable,
    estimate,
    moe,
    units)                             %>%
  mutate(a=abs(estimate-units) < moe)  -> acs_bg
print(acs_bg, n=50)
```

```{r get-acs-people}
acs1 <- get_acs(
  geography="block group", 
  variables="B01001_001",
  state="KS",
  county=c("91","103", "209"),
  geometry=FALSE, 
  cache_table=TRUE,
  year=2020)
glimpse(acs1)

acs2 <- get_acs(
  geography="block group", 
  variables="B01001_001",
  state="MO",
  county=c("37","47", "95", "165"),
  geometry=FALSE, 
  cache_table=TRUE,
  year=2020)
glimpse(acs2)

acs1                                   %>%
  bind_rows(acs2)                      %>%
  rename(bg_id=GEOID)                  %>%
  inner_join(bg_counts, by="bg_id")    %>%
  select(
    bg_id,
    variable,
    estimate,
    moe,
    people)                            %>%
  mutate(a=abs(estimate-people) < moe)  -> acs_bg
print(acs_bg, n=50)
```

<!---Key census links go here--->

[sumlev]: https://www2.census.gov/programs-surveys/decennial/rdo/about/2020-census-program/Phase3/SupportMaterials/FrequentSummaryLevels.pdf
