---
title: "Store block group information"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-14"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program will read in census block group files and store them in an RData file for later use.

### General information

The starting point for downloading most Census shapefiles is [here][c_general]. The ftp site is a bit more difficult to navigate, but leads you directly to the files that you want. 

It helps to know the FIPS (Federal Information Processing System) codes for Kansas (20) and Missouri (29).

Files with a postfix of "20" either in the filename or among some of the variable names indicates that the data or shapefile is from the 2020 Census. A postfix of "10" indicates that the data or shapefile is from the 2010 Census.

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Platte County, MO (29165)
Wyandotte County, KS (20209)

A census block group is a subdivision of a Census tract. These can be further subdivided into Census blocks.

Use the st_read function from the [sf package][r_sf] to read in the data. You may need to convert the coordinate reference system. That uses the st_crs function, also of the sf package. Census blocks are uniquely identified by a two digit state FIPS code, a three digit county FIPS code, a six digit tract code, and a single digit block group code. These are combined into a twelve digit code, GEOID.

```{r setup}
library(sf)
library(tidyverse)
path_name <- "../data"
```

### Kansas block groups

```{r read-ks}
co_list <- c(
  "091",
  "103",
  "209")
file_name <- "bg/tl_2021_20_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      %>%
  filter(COUNTYFP %in% co_list)        -> bg_ks
glimpse(bg_ks)
bg_ks                                  %>%
   ggplot()                             +
      geom_sf(aes())                    +
      ggtitle(file_name)
```



### Missouri block groups

```{r read-mo}
co_list <- c(
  "037",
  "047",
  "095",
  "165")
file_name <- "bg/tl_2021_29_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      %>%
  filter(COUNTYFP %in% co_list)         -> bg_mo
glimpse(bg_mo)
bg_mo                                  %>%
   ggplot()                             +
      geom_sf(aes())                    +
      ggtitle(file_name)
```

### Combine Kansas and Missouri

```{r combine}
bg_ks                                  %>%
  bind_rows(bg_mo)                      -> bg
```

### Rename variables

```{r rename}
bg                                     %>%
  mutate(bg_id=GEOID)                  %>%
  rename(bg_name=NAMELSAD)             %>%
  mutate(
    bg_area=
      as.numeric(st_area(bg)))         %>%
  select(
    bg_id,
    bg_name,
    bg_area,
    geometry)                           -> bg
glimpse(bg)
```

### Store as RData

```{r save}
save(bg, file="../data/bg.RData")
```
