---
title: "Store block group information"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-14"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program will read in census block group files and store them in an RData file for later use.

### General information

A census block group is a subdivision of a Census tract. These can be further subdivided into Census blocks. Census blocks are uniquely identified by a two digit state FIPS code, a three digit county FIPS code, a six digit tract code, and a single digit block group code. These are combined into a twelve digit code, GEOID.

The starting point for downloading most Census shapefiles is [here][c_general]. The ftp site is a bit more difficult to navigate, but leads you directly to the files that you want. 

There are files for census block groups for [Kansas][c_bg_ks] and [Missouri][c_bg_mo].
 
 [c_general]: https://www.census.gov/geographies/mapping-files/time-series/geo/tiger-line-file.html

[c_bg_ks]: https://www2.census.gov/geo/tiger/TIGER2021/BG/tl_2021_20_bg.zip
[c_bg_mo]: https://www2.census.gov/geo/tiger/TIGER2021/BG/tl_2021_29_bg.zip

The important counties for this program are

Cass County, MO (29037)
Clay County, MO (29047)
Jackson County, MO (29095)
Johnson County, KS (20091)
Leavenworth County, KS (20103)
Platte County, MO (29165)
Wyandotte County, KS (20209)



```{r setup}
library(sf)
library(tidyverse)
path_name <- "../data"
```

### Kansas block groups

```{r read-ks}
co_list <- c(
  "091",
  "103",
  "209")
file_name <- "bg/tl_2021_20_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      %>%
  filter(COUNTYFP %in% co_list)        -> bg_ks
glimpse(bg_ks)
bg_ks                                  %>%
   ggplot()                             +
      geom_sf(aes())                    +
      ggtitle(file_name)
```

### Missouri block groups

```{r read-mo}
co_list <- c(
  "037",
  "047",
  "095",
  "165")
file_name <- "bg/tl_2021_29_bg.shp"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  st_read(stringsAsFactors=FALSE)      %>%
  filter(COUNTYFP %in% co_list)         -> bg_mo
glimpse(bg_mo)
bg_mo                                  %>%
   ggplot()                             +
      geom_sf(aes())                    +
      ggtitle(file_name)
```

### Combine Kansas and Missouri

```{r combine-block-groups} 
bg_ks                                  %>%
  bind_rows(bg_mo)                     %>%
  rename(bg_id=GEOID)                  %>%
  rename(bg_name=NAMELSAD)              -> bg

bg                                     %>%
  mutate(
    bg_area=
      as.numeric(st_area(bg)))         %>%
  select(
    bg_id,
    bg_name,
    bg_area,
    geometry)                           -> bg

glimpse(bg)

bg                                     %>%
 ggplot()                               +
      geom_sf(aes())                    +
      ggtitle("mo-ks-tracts")  
```

### Store as RData

```{r save}
save(bg, file="../data/bg.RData")
```
