---
title: "Store redistricting information"
author: "Satya Golla and Steve Simon"
date: "Created 2022-03-08"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = "../results", output_format = "all") })  
---

This program will read in census redistricting files
which provide population and housing counts for
every census block in every state in the U.S.

### General information

The starting point for data and information about
the redistricting files is [here][c_red].

This program is based on pl_geohd_2020_dar.r, 
which is part of [this zip file][c_rcode].

[c_red]: https://www.census.gov/programs-surveys/decennial-census/about/rdo/summary-files.html
[c_rcode]: https://www2.census.gov/programs-surveys/decennial/rdo/about/2020-census-program/Phase3/SupportMaterials/2020PL_R_import_scripts.zip

An important contact at City Hall in Kansas City, Missouri is Steve Lebowski. He has done similar work every time new Census information is available.

### Setup

```{r setup}
library(sf)
library(tidyverse)
path_name <- "../data"
vnames <- c(
  "FILEID", 
  "STUSAB", 
  "SUMLEV", 
  "GEOVAR", 
  "GEOCOMP", 
  "CHARITER", 
  "CIFSN", 
  "LOGRECNO", 
  "GEOID", 
  "GEOCODE",
  "REGION",
  "DIVISION",
  "STATE",
  "STATENS",
  "COUNTY",
  "COUNTYCC",
  "COUNTYNS",
  "COUSUB",
  "COUSUBCC",
  "COUSUBNS",
  "SUBMCD",
  "SUBMCDCC",
  "SUBMCDNS",
  "ESTATE",
  "ESTATECC",
  "ESTATENS", 
  "CONCIT",
  "CONCITCC",
  "CONCITNS",
  "PLACE",
  "PLACECC",
  "PLACENS",
  "TRACT",
  "BLKGRP",
  "BLOCK", 
  "AIANHH",
  "AIHHTLI",
  "AIANHHFP",
  "AIANHHCC",
  "AIANHHNS",
  "AITS",
  "AITSFP",
  "AITSCC",
  "AITSNS",
  "TTRACT",
  "TBLKGRP",
  "ANRC",
  "ANRCCC",
  "ANRCNS",
  "CBSA",
  "MEMI",
  "CSA",
  "METDIV",
  "NECTA",
  "NMEMI",
  "CNECTA",
  "NECTADIV",
  "CBSAPCI",
  "NECTAPCI",
  "UA",
  "UATYPE",
  "UR",
  "CD116",
  "CD118",
  "CD119", 
  "CD120", 
  "CD121", 
  "SLDU18", 
  "SLDU22", 
  "SLDU24", 
  "SLDU26", 
  "SLDU28", 
  "SLDL18", 
  "SLDL22",
  "SLDL24",
  "SLDL26",
  "SLDL28",
  "VTD",
  "VTDI",
  "ZCTA",
  "SDELM",
  "SDSEC",
  "SDUNI",
  "PUMA",
  "AREALAND",
  "AREAWATR",
  "BASENAME",
  "NAME",
  "FUNCSTAT",
  "GCUNI",
  "POP100",
  "HU100",
  "INTPTLAT",
  "INTPTLON",
  "LSADC",
  "PARTFLAG",
  "UGA")
```

### Read Kansas redistricting files

```{r read-ks}
file_name <- "red/ksgeo2020.pl"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  read.delim(
    header=FALSE, 
    colClasses="character", 
    sep="|")                            -> red1
names(red1) <- vnames
red1                                  %>%
  filter(SUMLEV == "750")             %>%
  head
```

### Read Missouri redistricting files

```{r read-mo}
file_name <- "red/mogeo2020.pl"
path_name                              %>%
  paste(file_name, sep="/")            %>%
  read.delim(
    header=FALSE, 
    colClasses="character", 
    sep="|")                            -> red2
names(red2) <- vnames
red2                                  %>%
  filter(SUMLEV == "750")             %>%
  head
```

### Store as an RData file

Placing the file in an RData format will speed up future use of this file.

```{r store}
clist <- c(
  "20091",
  "20103",
  "20209",
  "29037",
  "29047",
  "29095",
  "29165")
red1                                   %>%
  bind_rows(red2)                      %>%
  filter(BLOCK != "")                  %>%
  mutate(GEOID=str_sub(GEOID, 10, 24)) %>%
  mutate(
    county_fips=
      str_sub(GEOID,  1,  5))          %>%
  filter(county_fips %in% clist)       %>%
  select(
    GEOID,
    POP100,
    HU100)                             %>%
  mutate(
    statefp=
      str_sub(GEOID,  1,  2))          %>%
  mutate(
    countyfp=
      str_sub(GEOID,  3,  5))          %>%
  mutate(
    tract=
      str_sub(GEOID,  6, 11))          %>%
  mutate(
    block=
      str_sub(GEOID, 12, 15))           -> red
glimpse(red)
save(red, file="../data/red.RData")
```
